\begin{eabstract}
 We give a computational interpretation of G\"odel--Dummett logic, based
 on the Curry--Howard isomorphism.  This leads to characterization
 of waitfree computation, which is a class of distributed
 computation.  Although both G\"odel--Dummett logic and waitfree computation are
 important players in their respective areas, their connection has not
 been known.  G\"odel--Dummett logic is an intermediate logic---a
 logic between classical and intuitionistic logics.
 Waitfree computation is a class of distributed computation where no process can
 wait for another.  Our contribution is defining a typed lambda calculus
 for G\"odel--Dummett logic, which can solve
 exactly the same problems solvable by waitfree computation.
 This is the first result characterizing shared memory consistency using
 a type system, marking an important step in
 an emerging area of research to be called ``implicit shared
 memory consistency.''

 In his Master's Thesis, the author sensed some semantic connection
 between G\"odel--Dummett logic and a degree of shared memory consistency
 called sequential consistency.
 However, we are heavily indebted to Arnon
 Avron's paper in 1991.  Our typed lambda calculus is based on Avron's
 hypersequents.
 Our problem comes from Avron's question
 ``it seems to us extremely important to determine the exact
 computational content of them [the logics with cut-elimination
 using hypersequents] and to develop corresponding $\lambda$-calculi.''
 To this twenty-year-old question, we give an answer for a typical
 example of G\"odel--Dummett logic.
 We show that the computational content of G\"odel--Dummett logic is
 waitfree computation.

 Technically, this research studies a typed lambda calculus.
 We show three theorems:
 no typed term can reduce infinitely often;
 no typed term can reduce to a term representing failure;
 and a typed term can solve a problem if and only if a waitfree protocol
 can solve the problem.
\end{eabstract}

\begin{jabstract}
 カリー・ハワード対応に基づいた，ゲーデル・ダメット論理の計算的解釈を与え
 る．すると，分散計算の問題クラスであるwaitfree計算を特徴づけられる．
 ゲーデル・ダメット論理もwaitfree計算も，それぞれの分野で重要であるのに，
 これらの関連は，これまで知られていなかった．ゲーデル・ダメット論理は中
 間論理(直観主義論理と古典論理の間の論理)の一つである．Waitfree計算は，プロセスが他の
 プロセスを待てないという制限で決まる，分散計算のクラスである．
 本研究の貢献は，ゲーデル・ダメット論理の型付きラムダ計算を定義して，
 ちょうどwaitfree計算で解ける問題だけを解けると示したことである．
 本研究は，
 共有メモリの一貫性を，型システムで特徴づけた初めての研究であって，
 興りつつある「非明示メモリ一貫性」と呼ぶべき研究分野の重要な一歩である．

 著者は修士論文で，ゲーデル・ダメット論理と共有メモリ一貫性の一種である
 逐次一貫性との意味論的な関連に気づいていた．
 しかし，本研究の手法も問題も
 Arnon Avronの1991年の論文によるところが大きい．
 まず，その論文でAvronが提案したハイパーシーケントが，
 本研究の型付きラムダ計算の基礎である．
 さらに，その論文でAvronが提起した問題
 「ハイパーシーケントでカット除去できる論理の計算的意味はなにか」
 が，本研究で探求する問題である．
 この二十年来の問題の典型例であるゲーデル・ダメット論理の場合に，解答を
 与える．ゲーデル・ダメット論理の計算的意味は，waitfree計算であると示す．

 技術的に，本研究は型付きラムダ計算の研究である．
 示した定理は，
 型が付いた項の簡約が無限に続かないことと，
 型が付いた項の簡約が失敗を表す項で終わらないことと，
 型が付いた項で解ける問題がwaitfreeに解ける問題と
 一致することの三つである．
\end{jabstract}
