\begin{acknowledge}
 I thank my adviser Masami Hagiya for his patience and some decisive
 comments at important moments.
 Even before composing this thesis, I became indebted to many of the
 juries and their colleagues.
 In an MLG (methematical logic group) meeting, Hiroakira Ono asked me
 about the constructive content of Dummett axiom, which I answer in
 Chapter~\ref{ch:lambda}.
 One of Naoki Kobayashi's colleagues Kohei Suenaga predicted that the
 content of Chapter~\ref{ch:lambda} would become a thesis.
 Ichiro Hasuo gave me indispensable advice on visiting the Netherland
 and studying there.
 I await valuable comments by Shinichi Honiden and Andrzej Murawski from
 practical and theoretical viewpoints.

 The work about hyper-lambda calculus is encouraged by feedbacks from
 ACAN (Algebraic and Coalgebraic Approaches to
 Non-Classical Logics Workshop) and OPLSS'11 participants,
 Pisa Proof Theory Workshop, FLOPS-2012
 and numerous other occasions.

 Grant-in-Aid for JSPS Fellows 23-6978 supported
 my second and third years of PhD, of which almost one year was spent
 in ILPS, the University of Amsterdam.
 The author thanks Maarten Marx, Maarten de Rijke and other members of ILPS,
 University of Amsterdam, for
 comfortable environments and stimulating discussions during my stay.
 Especially, working on a heavy problem for months with Alessandro
 Facchini and Evgeny Sherkhonov was a unique training.

 The author thanks Tadeusz Litak for encouragements and
 information on relevant research.
 At the University of Tokyo,
 discussions with Tatsuya Abe, Yoshihiko Kakutani and Masahiro Hamano
 nurtured my taste on formalistic, semantic and graphical approaches.
 Kazuhiko Yamamoto and some other people gave me an important advice
 that I concentrate on the thesis for a moment.
\end{acknowledge}
